Combinatorics of the geometry of Wilson loop diagrams

### 2019-02-07 General WLD combinatorics.tex
fixed the issue with the def of r_i in the last section, and at the same time fixed the scalar multiple

\begin{algorithm}\label{alg WLD to denom via GN}
Let $W = (\cP,n)$ be a Wilson loop diagram, and let $C(W)$ be the matrix of $W$ as defined in \eqref{C(W) dfn} (see Section~\ref{section background}).
\begin{itemize}
  \item For each $i \in [n]$, we construct a factor $r_i$ as follows:
    \begin{itemize}
      %\item Use Algorithm~\ref{alg:put GN on WLD} to obtain a bijection between the propagators and the $i-1$st Grassmann necklace element, $I_{i-1}$.  (By convention set $I_{-1}=I_n$.)  Write $I_{i-1}(p)$ for the vertex associated to propagator $p$ under this bijection.
      %\item Use Algorithm~\ref{alg:put GN on WLD} to obtain a bijection between the propagators and the $i$th Grassmann necklace element, $I_{i}$.  Write $I_{i}(p)$ for the vertex associated to propagator $p$ under this bijection.
      \item Let $S_i = \{p \in \cP \ | \ I_{i-1}(p) \neq I_i(p)\}$. (By convention, set $I_{-1} = I_n$.)
	  \item Write $\Delta_{I_i}$ for the determinant of the $k \times k$ minor of $C(W)$ with columns indexed by $I_i$.
      \item Let $r_i$ be $\Delta_{I_i}$ with all variables from rows associated to $p\not\in S_i$ set to $1$.
    \end{itemize}
  \item Define $R = \prod_{i=1}^n r_i$.
\end{itemize}
\end{algorithm}

As we take each $\Delta_{I_i}$ to simply be the determinant of a particular submatrix of $C(W)$, the sign of each $\Delta_{I_i}$ is well-defined.  The goal is to show that $R$ is equal to the denominator of the Wilson loop diagram as defined in Definition~\ref{def R(W)}.
%  However, we will not be interested in these signs as they only contribute an overall sign to $R$ and our main theorem in this section (Theorem~\ref{thm denom}) will simply be to show that $R$ is equal up to a $\mathbb{Q}$-multiple to the denominator of the Wilson loop diagram (see Definition~\ref{def R(W)}).

@@

Finally, we have $r_1 = (af-be)k$.  Putting everything together, we obtain
\[
R = (af-be)kgbclhdi(ej-fi)
\]
which is squarefree and contains all factors of $\prod_{i=1}^{n}\Delta_{I_i}$. If one were to construct the denominator $R(W)$ associated to this Wilson loop diagram as per Definition~\ref{def R(W)}, we would find that (up to integer multiples) we have $R(W) = R$.
\end{eg}

@@

\begin{proof}
\begin{enumerate}
    \item The nonzero entries of $C(W)$ are independent indeterminates and so every $i\times i$ minor of $C(W)$ is either homogeneous of degree $i$ or is $0$.  Thus each $\Delta_{I_i}$ is homogeneous.  Furthermore, each row contributes one factor to each term in the expansion of $\Delta_{I_i}$ so the result of setting the variables from a subset of rows to $1$ is still homogeneous.  Thus each $r_i$ is homogeneous.
    \item Using the expression for the determinant as a sum over permutations we see that $\Delta_{I_i}$ is a sum over bijections between $I_i$ and $\mathcal{P}$.  The nonzero terms in this sum are precisely those bijections such that each propagator is associated to one of its supporting vertices in $I_i$, since only those locations in $C(W)$ are nonzero.  Since the nonzero entries of $C(W)$ are independent there can be no cancellation between terms in this expansion.

@@

\item Consider $p\in S_i$, and note that $\Delta_{I_i}$ is homogeneous linear in the variables of the row corresponding to $p$.  By part (2), either exactly one variable in the row corresponding to $p$ appears in $\Delta_{I_i}$ and this variable is a factor of $\Delta_{I_i}$, or exactly two variables from the row corresponding to $p$ appear in $\Delta_{I_i}$ and they appear as part of a quadratic factor.  In the first case let the variable be $x$, then $x$ is a factor of both $r_i$ and $\Delta_{I_i}$ and is the only variable from this row in either polynomial.

@@

Thus $q\in S_i$ after all, and so again $f$ is a factor of $r_i$ as required.

@@

\begin{thm}\label{thm denom}
Given any admissible Wilson loop diagram $W$, let $\{I_1, \ldots I_n\}$ be the associated Grasmann necklace. Then the denominator of the integral, $R(W)$ (see Definition~\ref{def R(W)}), is a $\mathbb{Q}$-multiple of the radical of $\prod_{i=1}^n \Delta_{I_i}$, where $\Delta_{I_i}$ is the determinant of the $k \times k$ minor indicated by $I_i$.  Equivalently, $R(W)$ is a $\mathbb{Q}$-multiple of the $R$ of Algorithm~\ref{alg WLD to denom via GN}.
\end{thm}

\begin{proof}
The equivalence of the final statement of the theorem is due to Proposition~\ref{prop alg gives rad}.  It remains to prove that $R(W)$ is an integer multiple of the $R$ of Algorithm~\ref{alg WLD to denom via GN}.

To this end, first note that $R(W)$ and $R$ both have total degree $4|\mathcal{P}|$; the degree of $R(W)$ is immediate from the definition while that of $R$ follows from Lemma~\ref{vertex cyclic int lem}.  By Proposition~\ref{prop alg gives rad} every factor of $R$ is either a single variable or a quadratic factor coming from two propagators supported on a common edge.  The factors of each $R_e$ making up $R(W)$ in the notation of Definition~\ref{def R(W)} are all of this form and hence every factor of $R$ divides $R(W)$.
Finally, since $R$ is squarefree, this implies that $R(W)$ is a $\mathbb{Q}$-multiple of $R$.
\end{proof}

\note{Is there any point in working out the $\mathbb{Q}$-coefficient, or otherwise discussing the field?}

>>>>

\begin{algorithm}\label{alg WLD to denom via GN}
Let $W = (\cP,n)$ be a Wilson loop diagram, and let $C(W)$ be the matrix of $W$ as defined in \eqref{C(W) dfn} (see Section~\ref{section background}).
\begin{itemize}
  \item For each $i \in [n]$, we construct a factor $r_i$ as follows:
    \begin{itemize}
      %\item Use Algorithm~\ref{alg:put GN on WLD} to obtain a bijection between the propagators and the $i-1$st Grassmann necklace element, $I_{i-1}$.  (By convention set $I_{-1}=I_n$.)  Write $I_{i-1}(p)$ for the vertex associated to propagator $p$ under this bijection.
      %\item Use Algorithm~\ref{alg:put GN on WLD} to obtain a bijection between the propagators and the $i$th Grassmann necklace element, $I_{i}$.  Write $I_{i}(p)$ for the vertex associated to propagator $p$ under this bijection.
      \item Let $S_i = \{p \in \cP \ | \ I_{i-1}(p) \neq I_i(p)\}$. (By convention, set $I_{-1} = I_n$.)
	  %\item Write $\Delta_{I_i}$ for the determinant of the $k \times k$ minor of $C(W)$ with columns indexed by $I_i$.
      \item Let $r_i$ be the determinant of the $|S_i|\times |S_i|$ minor of $C(W)$ with rows indexed by $S_i$ and columns indexed by $I_i(S_i)$.  %$\Delta_{I_i}$ with all variables from rows associated to $p\not\in S_i$ set to $1$.
    \end{itemize}
  \item Define $R = \prod_{i=1}^n r_i$.
\end{itemize}
\end{algorithm}

As we discuss the algorithm it will also be useful to have the notation $\Delta_{I_i}$ for the determinant of the $k\times k$ minor of $C(W)$ with columned indexed by $I_i$.

As we take each $r_i$ to simply be the determinant of a particular submatrix of $C(W)$, the sign of each $r_i$ is well-defined.  The goal is to show that $R$ is equal to the denominator of the Wilson loop diagram as defined in Definition~\ref{def R(W)}.
%  However, we will not be interested in these signs as they only contribute an overall sign to $R$ and our main theorem in this section (Theorem~\ref{thm denom}) will simply be to show that $R$ is equal up to a $\mathbb{Q}$-multiple to the denominator of the Wilson loop diagram (see Definition~\ref{def R(W)}).

@@

Finally, we have $r_1 = (af-be)k$.  Putting everything together, we obtain
\[
R = (af-be)kgbclhdi(ej-fi)
\]
which is squarefree and contains all factors of $\prod_{i=1}^{n}\Delta_{I_i}$. If one were to construct the denominator $R(W)$ associated to this Wilson loop diagram as per Definition~\ref{def R(W)}, we would find that we have $R(W) = R$.
\end{eg}

@@

\begin{proof}
\begin{enumerate}
    \item The nonzero entries of $C(W)$ are independent indeterminates and so every $i\times i$ minor of $C(W)$ is either homogeneous of degree $i$ or is $0$.  Thus each $\Delta_{I_i}$ and each $r_i$ is homogeneous.  %Furthermore, each row contributes one factor to each term in the expansion of $\Delta_{I_i}$ so the result of setting the variables from a subset of rows to $1$ is still homogeneous.  Thus each $r_i$ is homogeneous.
    \item Using the expression for the determinant as a sum over permutations we see that $\Delta_{I_i}$ is a sum over bijections between $I_i$ and $\mathcal{P}$.  The nonzero terms in this sum are precisely those bijections such that each propagator is associated to one of its supporting vertices in $I_i$, since only those locations in $C(W)$ are nonzero.  Since the nonzero entries of $C(W)$ are independent there can be no cancellation between terms in this expansion.

@@

\item Consider $p\in S_i$, and note that $\Delta_{I_i}$ is homogeneous linear in the variables of the row corresponding to $p$.  By part (2), either exactly one variable in the row corresponding to $p$ appears in $\Delta_{I_i}$ and this variable is a factor of $\Delta_{I_i}$, or exactly two variables from the row corresponding to $p$ appear in $\Delta_{I_i}$ and they appear as part of a quadratic factor.  In the first case let the variable be $x$. Then $x$ is a factor of $\Delta_{I_i}$ and so in particular the monomial in $\Delta_{I_i}$ corresponding to the bijection between propagators and vertices of $I_i$ associates the column of $x$ to $p$.  Thus $x$ also appears in $r_i$ and since the matrix for $r_i$ is a minor of the matrix for $\Delta_{I_i}$ and every term in $\Delta_{I_i}$ involves $x$, we also have that every term in $r_i$ involves $x$ so $x$ is a factor of both $r_i$ and $\Delta_{I_i}$ and is the only variable from this row in either polynomial.

@@

Thus $q\in S_i$ after all, and so $f$ is a factor of $r_i$ as required.

@@

\begin{thm}\label{thm denom}
  Given any admissible Wilson loop diagram $W$, let $\{I_1, \ldots I_n\}$ be the associated Grasmann necklace. Then the denominator of the integral, $R(W)$ (see Definition~\ref{def R(W)}), the $R$ of Algorithm~\ref{alg WLD to denom via GN}, which is also up to scalar multiple the radical of $\prod_{i=1}^n \Delta_{I_i}$, where $\Delta_{I_i}$ is the determinant of the $k \times k$ minor indicated by $I_i$.  
\end{thm}

\note{Note that there is still a mention of scalar multiple in this theorem because it seems to me that the def of radical is only up to scalar multiple, but the part relating $R(W)$ to $R$ is exact with no scalar multiples.}

\begin{proof}
The equivalence up to scalar multiple of $R$ and the radical of $\prod_{i=1}^n \Delta_{I_i}$ is due to Proposition~\ref{prop alg gives rad}.  It remains to prove that $R(W)$ is the $R$ of Algorithm~\ref{alg WLD to denom via GN}.

To this end, first note that $R(W)$ and $R$ both have total degree $4|\mathcal{P}|$; the degree of $R(W)$ is immediate from the definition while that of $R$ follows from Lemma~\ref{vertex cyclic int lem}.  By Proposition~\ref{prop alg gives rad} every factor of $R$ is either a single variable or a quadratic factor coming from two propagators supported on a common edge.  The factors of each $R_e$ making up $R(W)$ in the notation of Definition~\ref{def R(W)} are all of this form and hence every factor of $R$ divides $R(W)$.
Finally, since $R$ is squarefree, this implies that $R(W)$ is a scalar multiple of $R$.

Finally then we need to check the scalar.  By Definition~\ref{def R(W)} each linear factor appears with coefficient $1$ and each $2\times 2$ deteriminant factor appears with the same sign as the determinant of the corresponding minor in $C(W)$.  Therefore $R=R(W)$.
\end{proof}
